---
title: "ML Happiness"
author: "Jasper"
date: "12/1/2020"
output: html_document
---

## libraries

```{r}

library(readr)
library(dplyr)
library(vroom)
library(janitor)
library(here)
library(Hmisc)
library(gapminder)
library(countrycode)
library(tidyverse)

```


## Load the data

### Happiness Data

```{r loading data}

data_2019 <- vroom::vroom(here::here("00_Data","2019.csv")) %>% 
  clean_names()

```

Adding additional info for matching purposes

```{r adding additional information}

data_2019_01 <- data_2019

data_2019_02 <- data_2019_01 %>% 
  mutate(continent = countrycode(sourcevar = data_2019_01[[2]],
                            origin = "country.name",
                            destination = "continent"),
         region = countrycode(sourcevar = data_2019_01[[2]],
                            origin = "country.name",
                            destination = "region"),
         iso2c = countrycode(sourcevar = data_2019_01[[2]],
                            origin = "country.name",
                            destination = "iso2c"),
         )

```


Imputing values for Kosovo:

continent == Europe
iso2c == XK, as used by the European Commission

```{r}

data_2019_03 <- data_2019_02 %>%
  mutate(iso2c = if_else(country_or_region=="Kosovo", 
                         "XK", 
                         iso2c) ,
         continent = if_else(country_or_region=="Kosovo",
                             "Europe",
                             continent)
  ) 

data_2019_03 %>% 
  group_by(iso2c) %>% 
  count() %>% 
  arrange(desc(n))

data_2019_03 %>% 
  filter(iso2c == "CY")

```

Northern Cyprus and Cyprus are listed as different countries.  
The iso standard however has them listed as one country.  
for our further analysis we will combine the two, taking the average

```{r cyprus, warning = FALSE}

#Aggreagating numeric values
cy_data <- aggregate(x = data_2019_03, by = list(data_2019_03$iso2c), FUN = mean) %>% 
  filter(Group.1 == "CY")

#replacing cyprus values with averages
data_2019_04 <- data_2019_03 %>% 
  mutate(score = if_else(country_or_region == "Cyprus", cy_data$score, score),
         gdp_per_capita = if_else(country_or_region == "Cyprus", cy_data$gdp_per_capita, gdp_per_capita),
         social_support = if_else(country_or_region == "Cyprus", cy_data$social_support, social_support),
         healthy_life_expectancy = if_else(country_or_region == "Cyprus",cy_data$healthy_life_expectancy, healthy_life_expectancy),
         freedom_to_make_life_choices = if_else(country_or_region == "Cyprus", cy_data$freedom_to_make_life_choices, freedom_to_make_life_choices),
         generosity = if_else(country_or_region == "Cyprus", cy_data$generosity, generosity),
         perceptions_of_corruption = if_else(country_or_region == "Cyprus", cy_data$perceptions_of_corruption, perceptions_of_corruption)
           ) %>% 
  #deleting northern cyprus
  filter(country_or_region != "Northern Cyprus")

```

```{r final data set happy}

happy_data_2019 <- data_2019_04

rm(data_2019, data_2019_01, data_2019_02, data_2019_03, data_2019_04, cy_data)

```

### Adding population data

Loading the population data

```{r loading pop data}
country_population <- read_csv("00_Data/country_population.csv", 
    skip = 3) %>% 
  clean_names() %>% dplyr::select(c(country_name, country_code, x2019))

```
cleaning the data and adding it to our data set

```{r adding pop data}
country_population <- country_population %>% 
  mutate(iso2c = countrycode(sourcevar = country_population[[1]],
       origin = "country.name",
       destination = "iso2c")) %>% 
  filter(!is.na(iso2c))

happy_data_2019_02 <- left_join(happy_data_2019, country_population, by = "iso2c")

happy_data_2019_02 <- happy_data_2019_02 %>% 
  mutate(x2019 = case_when(iso2c == "TW" ~ 23773876,
                           iso2c == "XK" ~ 1794248,
                           TRUE ~ x2019)) %>% 
  rename(population = x2019)

#https://data.worldbank.org/indicator/SP.POP.TOTL?locations=XK

happy_data_2019 <- happy_data_2019_02 %>% 
  dplyr::select(-c(country_name, country_code))
  
rm(happy_data_2019_02, country_population)
```


## Adding GDP Data 

loading data on gini index, the gap minder data set, employment and gni data

```{r loading wealth data}

gapminder_2019 <- vroom::vroom(here::here("00_Data","gapminder 2019.csv")) %>% 
  clean_names() %>% 
  rename(gdpPercap = x2019)


gini_2019 <- vroom::vroom(here::here("00_Data","gini.csv")) %>% 
  clean_names() %>% 
  rename(gini = x2019)


grossni_2019<- read.csv("00_Data/gni2019.csv") %>% 
  clean_names() %>% 
  rename(grossni = x2019) 


emp_2019<- read.csv("00_Data/employment.csv") %>% 
  clean_names() %>% 
  rename(emp_rate = x2019) 

```

cleaning the data and adding it to our data set

```{r}

#adding gapminder

gapminder_2019 <- gapminder_2019 %>% 
  mutate(iso2c = countrycode(sourcevar = country,
       origin = "country.name",
       destination = "iso2c")) %>% 
  filter(!is.na(iso2c))

gapminder_2019_merged <- merge(happy_data_2019, gapminder_2019, ) %>% 
  dplyr::select(-country)

#adding gini

gini_2019 <- gini_2019 %>% 
  mutate(iso2c = countrycode(sourcevar = country,
       origin = "country.name",
       destination = "iso2c")) %>% 
  filter(!is.na(iso2c)) %>% 
  dplyr::select(-c(x2018, country))

gapminder_2019_merged <- merge(gapminder_2019_merged, gini_2019)

#adding grossni

grossni_2019 <- grossni_2019 %>% 
  mutate(iso2c = countrycode(sourcevar = grossni_2019[[1]],
       origin = "country.name",
       destination = "iso2c")) %>% 
  filter(!is.na(iso2c)) %>% 
  dplyr::select(-c(country))

gapminder_2019_merged <- merge(gapminder_2019_merged, grossni_2019)

#adding employment

emp_2019 <- emp_2019 %>% 
  mutate(iso2c = countrycode(sourcevar = emp_2019[[1]],
       origin = "country.name",
       destination = "iso2c")) %>% 
  filter(!is.na(iso2c)) %>% 
  dplyr::select(-c(country))

happy_data_2019 <- merge(gapminder_2019_merged, emp_2019)


rm(gini_2019, gapminder_2019, gapminder_2019_merged)

```

imputing values for the 4 economic variables added above

```{r}

# adding data from IMF
# IMF: World Economic Outlook (WEO) Database, October 2020  Verified
# Link:https://knoema.com/IMFWEO2020Oct/imf-world-economic-outlook-weo-database-october-2020
data <- happy_data_2019 %>% 
  dplyr::mutate(
    gdpPercap = if_else(country_or_region == "Bhutan", as.numeric(11968), gdpPercap),
    gdpPercap = if_else(country_or_region == "Eritrea", as.numeric(1836), gdpPercap),
    gdpPercap = if_else(country_or_region == "Iran", as.numeric(12547), gdpPercap),
    gdpPercap = if_else(country_or_region == "Marshall Islands", as.numeric(3932), gdpPercap),
    gdpPercap = if_else(country_or_region == "Palau", as.numeric(15916), gdpPercap),
    # take 2018 value since no 2019....
    gdpPercap = if_else(country_or_region == "Palestine", as.numeric(5795), gdpPercap), 
    gdpPercap = if_else(country_or_region == "San Marino", as.numeric(63133), gdpPercap),
    gdpPercap = if_else(country_or_region == "South Sudan", as.numeric(862), gdpPercap),
    gdpPercap = if_else(country_or_region == "Tonga", as.numeric(6138), gdpPercap),
    gdpPercap = if_else(country_or_region == "Turkmenistan", as.numeric(16711), gdpPercap),
    gdpPercap = if_else(country_or_region == "Venezuela", as.numeric(7344), gdpPercap),
    gdpPercap = if_else(country_or_region == "Yemen", as.numeric(2057), gdpPercap),
    gdpPercap = if_else(country_or_region == "Papua New Guinea", as.numeric(4018), gdpPercap),
    gdpPercap = if_else(country_or_region == "Sudan", as.numeric(4140), gdpPercap),
    gdpPercap = if_else(country_or_region == "Angola", as.numeric(7384), gdpPercap),
    gdpPercap = if_else(country_or_region == "Solomon Islands", as.numeric(2590), gdpPercap),
    gdpPercap = if_else(country_or_region == "Guyana", as.numeric(13604), gdpPercap),
    gdpPercap = if_else(country_or_region == "Bahamas", as.numeric(39557), gdpPercap),
    gdpPercap = if_else(country_or_region == "Oman", as.numeric(33749), gdpPercap),
    gdpPercap = if_else(country_or_region == "Suriname", as.numeric(16768), gdpPercap),
    gdpPercap = if_else(country_or_region == "Fiji", as.numeric(14026), gdpPercap),
    gdpPercap = if_else(country_or_region == "Guinea-Bissau", as.numeric(2429), gdpPercap)
                                 )
data <- data %>% 
  dplyr::mutate(
    gdpPercap = if_else(iso2c == "AQ", as.numeric(1), gdpPercap),
    gdpPercap = if_else(iso2c == "GL", as.numeric(41800), gdpPercap),
    gdpPercap = if_else(iso2c == "CU", as.numeric(21016), gdpPercap),
    gdpPercap = if_else(iso2c == "KP", as.numeric(1700), gdpPercap),
    gdpPercap = if_else(iso2c == "SO" , as.numeric(900), gdpPercap),
    gdpPercap = if_else(iso2c == "SY" , as.numeric(2900), gdpPercap)
    )

happy_data_2019 <- data

rm(data, emp_2019, grossni_2019)

```



### Food

Loading the data

```{r}

food_supply <- vroom::vroom(here::here("00_Data","foodbalance.csv")) %>% 
  clean_names()

#deselecting rows
food_supply <- food_supply %>%
  dplyr::select(-c(domain_code, domain, area_code, element_code, item_code, year_code, flag, flag_description, unit))


```


Cleaning the data

```{r}
food_supply_tidy <- food_supply %>%
  pivot_wider(names_from="element", values_from="value") %>%
  mutate(category=case_when(item %in% c("Wheat and products", "Rice and products", "Barley and Products", "Maize and products", "Rye and products", "Oats", "Millet and products", "Sorghum and products", "Cereals, Other") ~ "Grains", item %in% c( "Potatoes and Products", "Sweet potatoes", "Yams", "Roots, Other") ~ "Starches", 
                            item %in% c("Sugar cane", "Sugar beet", "Sugar non-centrifugal", "Sugar (Raw Equivalent)", "Sweeteners, Other", "Honey")~"Sugar and Sweeteners", 
                            item %in% c("Beans", "Peas", "Pulses, Other and Products")~"Legumes", 
                            item %in% c("Nuts and products", "Soyabeans", "Groundnuts (Shelled Eq)", "Sunflower seed", "Rape and Mustardseed", "Cottonseed", "Sesame seed", "Coconuts - Incl Copra", "Palm kernels")~ "Nuts and seeds", 
         item %in% c("Oilcrops, Other", "Soyabean Oil", "Groundnut Oil", "Sunflowerseed Oil", "Rape and Mustard Oil", "Cottonseed Oil", "Palmkernel Oil", "Coconut Oil", "Sesameseed Oil", "Olive Oil", "Ricebran Oil", "Maize Germ Oil", "Oilcrops Oil, Other")~"Oils", 
         item %in% c("Olives (including preserved)", "Tomatoes and products", "Onions","Vegetables, Other", "Oranges, Mandarines", "Lemons, Limes, and products","Grapefruit and products", "Citrus, Other", "Bananas", "Plantains", "Apples and products", "Pineapples and products", "Dates", "Grapes and products (excl wine)", "Fruits, Other")~"Fruits and Vegetables", 
         item %in% c("Coffee and products", "Tea (including mate)")~"Coffee and Tea", 
         item %in% c("Cocoa Beans and products")~"Chocolate", 
         item %in% c("Pepper", "Pimento", "Cloves", "Spices, Other")~ "Spices", 
         item %in% c("Wine", "Beer", "Beverages, Fermented", "Beverages, Alcoholic", "Alcohol, Non-Food")~"Alcohol", item %in% c("Bovine meat", "Mutton & Goat Meat", "Pigmeat", "Poultry Meat", "Meat, Other", "Offals, Edible", "Meat, Aquatic Mammals")~ "Meat", 
         item %in% c("Butter, Ghee", "Cream", "Fats, Animals, Raw", "Eggs", "Milk")~"Eggs and Dairy", 
         item %in% c("Fish, Body Oil", "Fish, Liver Oil", "Freshwater Fish", "Demersal Fish", "Pelagic Fish", "Marine Fish, Other", "Crustaceans", "Cephalopods", "Molluscs, Other", "Aquatic Animals, Others", "Aquatic Plants")~ "Seafood", 
         item %in% c("Population", "Infant food", "Miscellaneous")~ "Other"))

#filter out just food categories 
food_supply_tidy <- food_supply_tidy %>%
  filter(category%in%c("Grains", "Meat", "Seafood", "Spices", "Fruits and Vegetables", "Oils", "Nuts and seeds", "Legumes", "Sugar and Sweeteners", "Starches", "Eggs and Dairy", "Chocolate"))%>%
  dplyr::select(-c("Total Population - Both sexes"))


#create iso2c column to match countries 
food_supply_tidy <- food_supply_tidy %>%
  mutate(iso2c = countrycode(sourcevar = food_supply_tidy[[1]],
                            origin = "country.name",
                            destination = "iso2c")) %>% 
  clean_names()%>%
  group_by(iso2c) %>%
  mutate(food_supply_kcal_capita_day = ifelse(is.na(food_supply_kcal_capita_day), 0, food_supply_kcal_capita_day))%>%
  summarise(food_supply_kcal_capita_day=sum(food_supply_kcal_capita_day)) 
  

```

Joining the data sets 

```{r}

#join two dataframes 
happy_data_2019 <- left_join(happy_data_2019, food_supply_tidy , on="iso2c")

rm( food_supply_tidy)
```

### Alcohol

loading the data

```{r}

alcohol_2019 <- vroom::vroom(here::here("00_Data","data_new.csv")) %>% 
  clean_names()

```

```{r}

alcohol_2019 <- alcohol_2019 %>% 
  mutate(iso2c = countrycode(sourcevar = country,
       origin = "country.name",
       destination = "iso2c")) %>% 
  filter(!is.na(iso2c)) %>% 
  dplyr::select(-c(score))
  

happy_data_2019 <- left_join(happy_data_2019, alcohol_2019 , by=c("iso2c" = "iso2c"))

rm(alcohol_2019)

```


### Gender

Loading gender data

```{r}

gender_2019 <- vroom::vroom(here::here("00_Data","Gender.csv")) %>% 
  clean_names()

```
Loading gender data

```{r}

gender_2019 <- gender_2019 %>% 
  filter(indicator == "Gender equality")

```
cleaning data and adding it to the data set

```{r}

gender_2019 <- gender_2019 %>% 
  dplyr::select(c(country_name, indicator_in_2018)) %>% 
  rename(gender_equ_indicator = indicator_in_2018) %>% 
    mutate(iso2c = countrycode(sourcevar = country_name,
       origin = "country.name",
       destination = "iso2c")) %>% 
  filter(!is.na(iso2c))
  

happy_data_2019 <- left_join(happy_data_2019, gender_2019 , by=c("iso2c" = "iso2c"))

```


```{r}
glimpse(happy_data_2019)
```

```{r}

library(showtext)
font_add(family = "Lato", regular = "Lato-Black.ttf")
showtext_auto()
```


# Exploring World Happiness Data 

## Map 

```{r}
#load map 
library(rnaturalearth)


world <- ne_countries(scale = "medium", returnclass = "sf") %>%
  filter(name != "Antarctica") %>%
  dplyr::select(c(geometry, iso_a2))

happy_world <- left_join(world, happy_data_2019, by=c("iso_a2" = "iso2c")) %>%
  clean_names()


ggplot(happy_world)+
  geom_sf(aes(fill=score),     
          show.legend = TRUE, color="white", size=0.2)+    # no legend
  scale_fill_gradient(low = "#DFF5FD", high = "#31489F", na.value = "#DCDCDC")+
  coord_sf(datum = NA) +
  theme_minimal()+
  labs(title="How happy are we?", subtitle= "Happiness score by country", fill= "Happiness Score", caption = "Source: World Happiness Report 2019")  + 
    theme(plot.title = element_text(face = "bold", size=12, family="Lato"),
        plot.caption = element_text(size=7, color="gray"),
        plot.subtitle = element_text(size=9),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "gray"), 
        axis.text.x = element_text(size=7),
        axis.text.y = element_text(size=9),
        axis.title = element_text(size=9, vjust = -1), 
        legend.title = element_text (size=9), 
        legend.position = "bottom")
  

```


## Ridge Plot 

```{r}
library(ggridges)
library(ggplot2)
#group by continent 
happy_data_2019_labels <- happy_data_2019 %>% 
  group_by(continent) %>% 
  summarise(countries = n()) %>% 
  arrange(countries)

continents_ord <-c("Europe","Asia","Americas","Africa", "Oceania")

ridge_labels <- c( "Europe\n(n=40)","Asia\n(n=45)","Americas\n(n=23)","Africa\n(n=45)", "Oceania\n(n=2)")


#mapping distributions across continents 

ggplot(happy_data_2019, aes(x = score, y = factor(continent, levels = continents_ord), 
                            fill = continent, height = (stat(count)))
       ) +
    geom_density_ridges(alpha=0.6, stat="binline", bins=20, scale = 0.95) +
    theme_ridges() +
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8)
    ) +
    xlab("Happiness Score") +
    ylab("") +
  labs(title="Africa is the most unhappy continent", subtitle="Distribution of happiness across continents", caption = "Source: World Happiness Report 2019") +
  scale_y_discrete(labels= ridge_labels)+
  scale_fill_manual(values = c("#4E67C8", "#B4DCFA", "#FDB519", "#5DCEAF", "#F14124"))+ 
   theme(plot.title = element_text(face = "bold", size=12, family="Lato"),
        plot.caption = element_text(size=7, color="gray"),
        plot.subtitle = element_text(size=9),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "gray"), 
        axis.text.x = element_text(size=7),
        axis.text.y = element_text(size=9),
        axis.title = element_text(size=9, vjust = -1),
        legend.position = "none")
  
  #geom_label(data = happy_data_2019_labels, aes(label = countries, y = continent)) +
  NULL
```

## Top and Bottom 

```{r}
happy_data_2019 %>% 
  mutate(score_norm = score - mean(score)) %>% 
  filter( min_rank(desc(score_norm)) <= 5 | 
                min_rank(score_norm) <= 5 ) %>% 
  mutate(top_bottom = if_else(score_norm>=0, "top", "bottom"),
         top_label = as.numeric(ifelse(top_bottom == "top", score_norm, "")),
         bottom_label = as.numeric(ifelse(top_bottom == "bottom", score_norm, ""))) %>% 
  add_row(score_norm = 0, country_or_region = "") %>% 
  
  
  ggplot(aes(x = score_norm, y = reorder(country_or_region,score_norm))) +
  geom_col(aes(fill = top_bottom)) +
  theme_minimal() +
  labs(y = "", title= "The top countries in terms of happiness are all in Northern \nEurope", subtitle="Difference of the top five happiness scores and bottom five happiness scores \ncompared to the average", x = "Average score = 5.45") +
  theme(legend.position = "none") +
  scale_fill_manual(values=c("#B4DCFA", "#FDB519"))+ 
  geom_text(aes(label=round(top_label,1), fontface=2), hjust=1.5, size=3) +
  geom_text(aes(label=round(bottom_label,1), fontface=2),hjust=-0.8, size=3) +
  geom_vline(xintercept = 0, color="gray") + 
theme(plot.title = element_text(face = "bold", size=12, family="Lato"),
        plot.subtitle = element_text(size=9),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "gray"), 
        axis.text.x = element_blank(),
        axis.text.y = element_text(size=9),
        axis.title.x.bottom = element_text(size=9, vjust = -1, family= "Lato"), 
        legend.title = element_text (size=9))


```

# Does money buy happiness? 

```{r}

library(dplyr)
library(vroom)
library(janitor)
library(here)
library(Hmisc)
library(gapminder)
library(countrycode)
library(rworldmap)
library(tidyverse)
library(rasterVis)
#library(plot3D)
library(quantreg)
library(corrplot)
library(RColorBrewer)
library(gridExtra)

```

## Faceted by Gini Index 

```{r combined plot}

# Break up into 3 groups according to wealth equality
happy_data_2019_gap <- happy_data_2019 %>% 
mutate(
    gini_category = case_when(
      gini <= 33 ~ "Equal socities \n(Gini < 0.33)",
      gini <= 40 ~ "Moderate socities \n(Gini = 0.33-0.40)",
      gini > 40 ~ "Unequal socities \n(Gini > 0.40)"))

# Plot
ggplot(happy_data_2019_gap, aes(y=score, x=log(gdpPercap), color=gini_category))+
  geom_point(alpha=0.5) +
  labs(title="Does money buy happiness?", subtitle = "Relationship of GDP per capita (PPP) and Happiness Score in different societal structures", x="GDP per capita, PPP (log $)", y="Happiness Score", caption ="Source: Gapminder & WHR 2019", color="Gini Coefficient")   +
  theme_minimal() + 
  facet_wrap(~gini_category) +
geom_smooth(method="lm")+ 
  scale_color_manual(values = c("#FDB519", "#4E67C8", "#5DCEAF"))+ 
  theme(plot.title = element_text(face = "bold", size=12, family="Lato"),
        plot.caption = element_text(size=7, color="gray"),
        plot.subtitle = element_text(size=9),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "gray"), 
        axis.text.x = element_text(size=7),
        axis.title = element_text(size=9, vjust = -1),
        legend.position = "none")



```

## Correlation Matrix 

```{r combined plot, warning=FALSE}
library(GGally)

#Select library
happy_data_corr<- happy_data_2019%>%
  dplyr::select(score, gdpPercap, gini, grossni, emp_rate, continent) 

#GGpair
ggpairs(happy_data_corr, columns = 1:5, ggplot2::aes(fill=continent, alpha=0.6), alpha=0.4) + theme_minimal()+ 
labs(title="Economic factors as an important foundation to happiness", subtitle = "Correlation between various economic indicators and happiness", caption ="Source: Gapminder & WHR 2019", color="Continent") +
    scale_fill_manual(values = c("#4E67C8", "#B4DCFA", "#FDB519", "#5DCEAF", "#F14124"))+ 
   theme(plot.title = element_text(face = "bold", size=12, family="Lato"),
        plot.caption = element_text(size=7, color="gray"),
        plot.subtitle = element_text(size=9),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "gray"), 
        axis.text.x = element_text(size=7),
        axis.title = element_text(size=9, vjust = -1), 
        legend.title = element_text(size=9, face="bold"), 
        legend.position="bottom")
  


```

## Map 

```{r plotting wealth map}

data <- happy_data_2019

# Create world map object
world<-map_data("world")

# Store country names
colnames(world)[5] <- "Country"

# Get ISO2C code
world <- world %>% 
mutate(iso2c = countrycode(sourcevar = world[[5]],
                          origin = "country.name",
                          destination = "iso2c"),
       ) %>%
  filter(Country != "Antarctica")


# Join gapminder to world
data <- left_join(world, happy_data_2019, by = c("iso2c" = "iso2c")) # ISO-code...

# Plot gdpPercap
ggplot()  +
  geom_polygon(data=data, aes(x = long, y = lat, group = group,fill= gdpPercap)) + 
  coord_equal() +scale_fill_gradientn(breaks=c(3,5,7,9), colours=rainbow(8)) +
  labs(title="Where are the riches?", subtitle="Distribution of GDP per capita across the globe") +
  xlab("") + ylab("") + guides(shape=FALSE) + labs(fill="GDP Per Capita")+
scale_fill_gradient(low = "#FCD9D3", high = "#F14124", na.value = "#DCDCDC") +
   theme_minimal() + 
  theme(plot.title = element_text(face = "bold", size=12, family="Lato"),
        plot.caption = element_text(size=7, color="gray"),
        plot.subtitle = element_text(size=9),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_blank(), 
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title = element_text(size=9, vjust = -1), 
        legend.title = element_text (size=9), 
        legend.position = "bottom", 
        legend.text = element_text(size = 6, color="black"))
```



# Fooooooood!!! 


## Food scatterplot 

```{r}
 
#food vs. happiness
happy_data_2019 %>%
  filter(iso2c != "CN")%>%
  filter(continent!="NA") %>% #remove outlier China cause otherwise it messes with the scale 
  ggplot(aes(x=food_supply_kcal_capita_day, y=score))+
  geom_point(shape=21, alpha=0.8, aes(fill=continent)) +
  geom_smooth(method="lm", color = "black", size = 0.5, alpha=0.2) + 
  theme_minimal()+ 
  labs(title="Food makes people happy", subtitle="Relationship between food supply (kcal/capita/day) and happiness score", x="Food Supply (kcal/capita/day)", y="Happiness Score", fill="Continent", caption = "World Happiness Report 2019, Food and Agriculture Organization of the United Nations")+ 
  scale_fill_manual(values = c("#4E67C8", "#B4DCFA", "#FDB519", "#5DCEAF", "#F14124"))+ 
   theme(plot.title = element_text(face = "bold", size=12, family="Lato"),
        plot.caption = element_text(size=7, color="gray"),
        plot.subtitle = element_text(size=9),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "gray"), 
        axis.text.x = element_text(size=7),
        axis.title = element_text(size=9, vjust = -1), 
        legend.title = element_text(size=9, face="bold"), 
        legend.position="bottom")
```

## Fishies 



```{r}
food_supply_tidy <- food_supply %>%
  pivot_wider(names_from="element", values_from="value") %>%
  mutate(category=case_when(item %in% c("Wheat and products", "Rice and products", "Barley and Products", "Maize and products", "Rye and products", "Oats", "Millet and products", "Sorghum and products", "Cereals, Other") ~ "Grains", item %in% c( "Potatoes and Products", "Sweet potatoes", "Yams", "Roots, Other") ~ "Starches", 
                            item %in% c("Sugar cane", "Sugar beet", "Sugar non-centrifugal", "Sugar (Raw Equivalent)", "Sweeteners, Other", "Honey")~"Sugar and Sweeteners", 
                            item %in% c("Beans", "Peas", "Pulses, Other and Products")~"Legumes", 
                            item %in% c("Nuts and products", "Soyabeans", "Groundnuts (Shelled Eq)", "Sunflower seed", "Rape and Mustardseed", "Cottonseed", "Sesame seed", "Coconuts - Incl Copra", "Palm kernels")~ "Nuts and seeds", 
         item %in% c("Oilcrops, Other", "Soyabean Oil", "Groundnut Oil", "Sunflowerseed Oil", "Rape and Mustard Oil", "Cottonseed Oil", "Palmkernel Oil", "Coconut Oil", "Sesameseed Oil", "Olive Oil", "Ricebran Oil", "Maize Germ Oil", "Oilcrops Oil, Other")~"Oils", 
         item %in% c("Olives (including preserved)", "Tomatoes and products", "Onions","Vegetables, Other", "Oranges, Mandarines", "Lemons, Limes, and products","Grapefruit and products", "Citrus, Other", "Bananas", "Plantains", "Apples and products", "Pineapples and products", "Dates", "Grapes and products (excl wine)", "Fruits, Other")~"Fruits and Vegetables", 
         item %in% c("Coffee and products", "Tea (including mate)")~"Coffee and Tea", 
         item %in% c("Cocoa Beans and products")~"Chocolate", 
         item %in% c("Pepper", "Pimento", "Cloves", "Spices, Other")~ "Spices", 
         item %in% c("Wine", "Beer", "Beverages, Fermented", "Beverages, Alcoholic", "Alcohol, Non-Food")~"Alcohol", item %in% c("Bovine meat", "Mutton & Goat Meat", "Pigmeat", "Poultry Meat", "Meat, Other", "Offals, Edible", "Meat, Aquatic Mammals")~ "Meat", 
         item %in% c("Butter, Ghee", "Cream", "Fats, Animals, Raw", "Eggs", "Milk")~"Eggs and Dairy", 
         item %in% c("Fish, Body Oil", "Fish, Liver Oil", "Freshwater Fish", "Demersal Fish", "Pelagic Fish", "Marine Fish, Other", "Crustaceans", "Cephalopods", "Molluscs, Other", "Aquatic Animals, Others", "Aquatic Plants")~ "Seafood", 
         item %in% c("Population", "Infant food", "Miscellaneous")~ "Other"))
```



```{r}
#filter out just food categories 
food_supply_tidy <- food_supply_tidy %>%
  filter(category%in%c("Grains", "Meat", "Seafood", "Spices", "Fruits and Vegetables", "Oils", "Nuts and seeds", "Legumes", "Sugar and Sweeteners", "Starches", "Eggs and Dairy", "Chocolate"))%>%
  dplyr::select(-c("Total Population - Both sexes"))

```


```{r}
#merge with happiness dataframe

#create iso2c column to match countries 
food_supply_tidy <- food_supply_tidy %>%
  mutate(iso2c = countrycode(sourcevar = food_supply_tidy[[1]],
                            origin = "country.name",
                            destination = "iso2c"))


#join two dataframes 
happy_food <- left_join(food_supply_tidy, happy_data_2019, on="iso2c")
```



```{r}
happy_food %>%
 clean_names()%>%
  filter(score!="NA")%>%
  summarise(average = mean(score))

#average happiness score = 5.45 


happy_food_sea <- happy_food %>%
  clean_names() %>%
  group_by(iso2c, continent, score, category, country_or_region) %>%
  mutate(food_supply_kcal_capita_day = ifelse(is.na(food_supply_kcal_capita_day), 0, food_supply_kcal_capita_day))%>%
  summarise(food_supply_kcal_capita_day=sum(food_supply_kcal_capita_day)) %>%
  filter(category=="Seafood", score!= "NA", food_supply_kcal_capita_day>=98) %>%
  arrange(desc(food_supply_kcal_capita_day))

#plotting top 10 fish consuming countries 
happy_food_sea %>%
  mutate(above_avg = case_when(score >= 5.45 ~ TRUE, score < 5.45 ~ FALSE))%>%
  ggplot(aes(x=reorder(country_or_region,-food_supply_kcal_capita_day), y=score)) + 
  geom_col(aes(fill=above_avg))+ 
  geom_hline(yintercept=5.45, linetype="dashed", color = "#B4DCFA") + 
  annotate("text", x = 8.5, y=5.8, label = "Average global happiness score = 5.45", size=3, family="Lato")+ 
  theme_minimal() + 
  scale_fill_manual(values = c( 'gray', '#4E67C8'))+ 
  labs(title="Six of the top ten seafood consuming countries have a \nhappiness score above average", x= "Top 10 seafood consuming countries (kcal/capita/day)", y= "Happiness score", caption = "World Happiness Report 2019, Food and Agriculture Organization of the United Nations")+ 
   theme(plot.title = element_text(face = "bold", size=12, family="Lato"),
        plot.caption = element_text(size=7, color="gray"),
        plot.subtitle = element_text(size=9),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "gray"), 
        axis.text.x = element_text(size=7),
        axis.title = element_text(size=9, vjust = -1),
        legend.position = "none")
  


```


# Alcohol 

```{r}
#bubble plot 
happy_data_alc<-happy_data_2019 %>% 
  group_by(country) %>% 
  mutate(most_loved=if_else(beer_servings>wine_servings & beer_servings>spirit_servings,'Beer',
                            if_else(wine_servings>beer_servings & wine_servings>spirit_servings,'Wine','Spirits')))
data <- happy_data_alc %>% 
  mutate(most_loved=if_else(beer_servings==0 & wine_servings==0 & spirit_servings==0,'None',most_loved))



ggplot(data,aes(x=total_litres_of_pure_alcohol,y=score,size=gdp_per_capita.y,color=most_loved))+
  geom_point(alpha=0.5)+
  theme_minimal()+
  labs(title='In happier countries people drink more, but they are also richer!',subtitle='Beer is the most consumed drink overall',x='Litres of alcohol per person',y='Happiness Score',caption='Source: WHR & FivethirtyEight')+
   theme_minimal() + 
  scale_color_manual(values=c("#FF8021", "#5ECCF3", "#31489F", "#5DCEAF"))+ 
   theme(plot.title = element_text(face = "bold", size=12, family="Lato"),
        plot.caption = element_text(size=7, color="gray", hjust=1),
        plot.subtitle = element_text(size=9),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "gray"), 
        axis.text.x = element_text(size=7),
        axis.title = element_text(size=9, vjust = -1), 
        legend.title= element_text(face="bold", size=8))+ 
  labs(size="GDP Per Capita",color='Most Consumed')
```

```{r}
#boxplot
data2<- data %>% 
  filter(most_loved %in% c('Beer','Wine','Spirits')) %>% 
  group_by(most_loved)


ggplot(data2,aes(x=reorder(most_loved,-score),y=score))+
  geom_boxplot(alpha=0.6, aes(fill=most_loved))+
  theme_minimal()+
  labs(title='Countries that consume wine the most are the happiest', subtitle='Box and whisker plot of countries that consume mostly wine, beer, or spirits.',x='Most Consumed',y='Score',caption='Source: WHR & FivethirtyEight')+
   theme_minimal() + 
  scale_fill_manual(values=c("#FF8021", "#31489F", "#5DCEAF"))+ 
   theme(plot.title = element_text(face = "bold", size=12, family="Lato"),
        plot.caption = element_text(size=7, color="gray", hjust=1),
        plot.subtitle = element_text(size=9),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "gray"), 
        axis.text.x = element_text(size=7),
        axis.title = element_text(size=9, vjust = -1), 
        legend.position="none")
```


# Gender Equality 

```{r}
library(RColorBrewer)
display.brewer.pal(10,"RdYlGn")
library(rnaturalearth)
library(countrycode)
library(sf)

```


# Map 

```{r Gender Equality map}
earth <- ne_countries(scale = 50, returnclass = "sf") %>% 
  dplyr::select(admin, geometry, iso_a2)

equality <- left_join(earth, happy_data_2019, by = c("iso_a2" = "iso2c")) %>% 
  sf::st_as_sf() %>%
  filter(admin != "Antarctica")



equality %>%
  ggplot() +
  geom_sf(aes(fill = gender_equ_indicator), size = 0.1, color = "White") +
  coord_sf(crs = "+proj=longlat +ellps=WGS84")  + #Mollweide projection
scale_fill_gradient(low = "#DFF5EF", high = "#5DCEAF", na.value = "#DCDCDC")+ 
  labs(title = "Gender equality across the world", fill = "Gender equality index") + 
  theme_minimal() + 
   theme(plot.title = element_text(face = "bold", size=12, family="Lato"),
        plot.caption = element_text(size=7, color="gray"),
        plot.subtitle = element_text(size=9),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_blank(), 
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title = element_text(size=9, vjust = -1), 
        legend.title = element_text (size=9), 
        legend.position = "bottom")
  


```

# Bubble Plot 

```{r}
library(ggplot2)
library(ggExtra)
library(MASS)
library(viridis)
library(ggrepel)
library(dichromat)

#happy_data_2019_02 <- happy_data_2019_02 #%>% 
 # select(country_or_region, region, score)

index_1 <- happy_data_2019 %>%
  filter(!is.na(region))
#View(index_1)
 
  ggplot(index_1, aes(x=gender_equ_indicator,
                    y= score, fill = continent, size = population)) +
  geom_point(alpha = 0.5, shape = 21, color = "black") +
    
    xlab("Gender Equality Index") +
    ylab("Happiness Index") +
   # facet_wrap(~region) +
 
    labs(title = "More equal countries tend to be happier", subtitle = "Gender equality index vs. happiness score, \nby continent and population size", fill = "Continent") + 
    
  theme_minimal() +
    
  scale_size_continuous(breaks = c(5e+08, 1e+09), labels = c("500m", "1000m"), range = c(2, 22), name="Country Population Size")+ # Legend name + Size
    
  theme(legend.position = "right", legend.title = element_text(face = "bold", size = 9),
  legend.text = element_text(size = 8)) + # Legend position
  
    scale_fill_manual(values = c("#4E67C8", "#B4DCFA", "#FDB519", "#5DCEAF", "#F14124")) +# Color palette with legends shown
    #scale_fill_discrete(name = "Continent", labels = c("Europe", "Oceania", "Americas", "Asia", "Africa")) + 
    theme(plot.title = element_text(face = "bold", size=12, family="Lato"),
        plot.caption = element_text(size=7, color="gray"),
        plot.subtitle = element_text(size=9),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "gray"), 
        axis.text.x = element_text(size=7),
        axis.title = element_text(size=9, vjust = -1))
```





# ML model 

```{r}
happy_analysis <- happy_data_2019 %>% 
  dplyr::select(-c(country_name,most_loved,spirit_servings,wine_servings,beer_servings, gdp_per_capita.y
))
happy_analysis %>% 
  dplyr::select(c(iso2c, gender_equ_indicator, ))
```

```{r}
# Fitting Decision Tree Regression to the dataset
library(rpart)
library(caTools)
library(rpart.plot)
library(extrafont)
#setting seed
set.seed(1)
dataset <- happy_data_2019 %>% 
  #mutating gdp to enhance readability
  mutate(gdpPercap_k = gdpPercap / 1000) %>% 
  #keeping only columns we want our tree model to use
  dplyr::select(c(score, gini, gdpPercap_k, food_supply_kcal_capita_day, total_litres_of_pure_alcohol, gender_equ_indicator))
# splitting data in test and training
split = sample.split(dataset$score, SplitRatio = 0.7)
training_set = subset(dataset, split == TRUE)
test_set = subset(dataset, split == FALSE)
#training tree model
regressor_dt = rpart(formula = score ~ .,
                  data = training_set,
                  control = rpart.control(minsplit = 10))
# Predicting a new result with Decision Tree Regression
y_pred_dt = predict(regressor_dt, newdata = test_set)
# combining actual and prediction scores for later visualization
Pred_Actual_dt <- as.data.frame(cbind(Prediction = y_pred_dt, Actual = test_set$score))
# plotting rules
rpart.plot(regressor_dt, main = "Results of the Decision Tree Model", box.palette = c("#B4DCFA","#5ECCF3","#4E67C8"), tweak = 1.1, shadow.col = "gray",family = "Lato")
```


```{r}
# picking a country to highlight in later plots
# adding country data to testing set
x <- happy_data_2019[row.names(Pred_Actual_dt),]
# choosing GB as example
Pred_Actual_dt <- Pred_Actual_dt %>% 
  mutate(format = x$country_or_region == "United Kingdom")
#setting label to be siplayed in plot
label = "Despite limited prediction \npower for most countries,\n Brits seem to follow \nour tree rules!"
# plotting reuslts
ggplot(Pred_Actual_dt, aes(Actual, Prediction)) +
  geom_point(color = "#B4DCFA", size = 2)  +
  theme_minimal() + 
  geom_abline() +
  labs(title = "Actual Happiness diverts from our predictions, indicating bad model performance", subtitle= "Model predicitions vs. actual happiness score",x = "Actual happiness score",
       y = "Predicted happiness score") +
  scale_color_manual(values = c("#B4DCFA", "#FDB519")) +
  #adding arrow
  geom_curve(
    data = data.frame(x = 7.1, y = 5.5, xend = 7.1, yend = 7.029143),
    mapping = aes(x = x, y = y, xend = xend, yend = yend),
    colour = "grey15",
    size = 0.5,
    curvature = 0.25,
    arrow = arrow(length = unit(2, "mm"), type = "closed"),
    inherit.aes = FALSE ) +
  # adding text field
  geom_text(
    data = data.frame(x = 7.1, y = 5, label = label),
    aes(x = x, y = y, label = label),
    colour="#FDB519",
    family="Comic Sans MS",
    hjust = 0.5,
    lineheight = .8,
    inherit.aes = FALSE,
  ) +  theme(plot.title = element_text(face = "bold", size=12, family="Lato"),
        plot.caption = element_text(size=7, color="gray"),
        plot.subtitle = element_text(size=9),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "gray"), 
        axis.text.x = element_text(size=7),
        axis.title = element_text(size=9, vjust = -1),
        legend.position = "none") + 
  geom_point(data=Pred_Actual_dt %>% filter(format==TRUE),
             aes(x=Actual, y=Prediction), size = 4, color = "#FDB519")
  
```




